---
title: "Statistical Inference"
author: "Sergei Titov"
subtitle: Peer Assessment Part 1
output: pdf_document
---

## Overview
This part of Statistical Inference project consider with the exponential distribution and the Central Limit Theorem. At this report we will:

- Generate a sample of a thousand simulations of averages of 40 exponentially distributed randoms
- Show the sample mean and compare it to the theoretical mean of the distribution
- Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution
- Show that the distribution is approximately normal

## Simulation
Set the default parameters.
```{r, echo = T}
lambda <- 0.2
n <- 40
nosim <- 1000
```

Fast look on original exponential distribution.
```{r, fig.height = 5, fig.width = 6}
library ("ggplot2")
g <- ggplot (data.frame (x = rexp (nosim, lambda)), aes(x = x))
g <- g + geom_histogram (alpha = .20, binwidth = .3, 
                         colour = "black", aes(y = ..density..))
g
```

Create a matrix with 1000 rows, 40 observations per row and calculate row means.
```{r, echo = T}
set.seed (1234)
data <- matrix (rexp (nosim * n, lambda), nrow = nosim)
data <- rowMeans(data)
```

Plotting data.
```{r, fig.height = 5, fig.width = 6}
g <- ggplot (data.frame (x = data), aes(x = x))
g <- g + geom_histogram (alpha = .20, binwidth = .15, 
                         colour = "black", aes (y = ..density..))
g
```

## Sample vs. theoretical parameters
```{r, echo = T}
sample_mean <- mean (data)
sample_var <- var (data)
sample_sd  <- sd (data)
```

The sample mean equals `r round (sample_mean, 3)` and  theoretical mean of the distribution $1/lambda = 1/0.2 = 5$  
The variance of sample means is `r round (sample_var, 3)`, where the theoretical variance of the distribution is 
$(1/lambda) / n = 5 / 40 = 0.625$  
The sample mean equals `r round (sample_sd, 3)` and  theoretical standard deviation equals 
$(1 / lambda) / \sqrt {n} = (1 / 0.2) / \sqrt {40} =$ `r round (5 / sqrt (40), 3)`

##Comparison with Normal distribution

To compare the sample distribution with normal distribution plot density functions.
```{r, fig.height = 5, fig.width = 8}
g <- ggplot (data.frame (x = data), aes(x = x)) +
geom_histogram (alpha = .2, fill = "black", binwidth = .15, 
                colour = "black", aes (y = ..density..)) + 
geom_density (fill = NA, aes (colour = "sample")) + 
stat_function (fun = dnorm, args = list (mean = 1/lambda, sd = 5/sqrt (n)), 
               aes (colour = "theoretical")) +
scale_colour_manual (name = "Density", 
                     values = c ("sample" = "black", "theoretical" = "red"))
g
```

Q-Q Plot.
```{r, fig.height = 5, fig.width = 5}
ggplot(data = as.data.frame(qqnorm (data ,plot=F)), mapping = aes(x=x, y=y)) + 
    geom_point() + geom_smooth(method="lm", se=FALSE)
```

Due to the central limit theorem (CLT), the distribution of averages of 40 exponentials is very close to a normal distribution.

